---
title: "Amoree Sales Data analysis 2022"
output:
  html_document: default
  pdf_document: default
date: "2023-11-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Markdown for 2022 Amoree Sales data analysis
  
  

### Install basic packages

```{r}
library(tidyverse)
library(magrittr)
library(readxl)
library(dplyr) 
library(tibble)  
```


### Import 2022 data from excel file


```{r}
Amoree2022 <- read_excel("C:\\Users\\91914\\Documents\\Data analysys\\Project Amoree Gifting\\Amoree Gifting Sales.xlsx",sheet = "2022")


Amoree2022 <- data.frame(Amoree2022)
```


### Rename columns  

```{r}
Amoree2022_v2 <- Amoree2022 %>% 
  rename(Sl_No=Sl.No.,Customer=Client.Name.,Price=Sale.Split....,Date=Delivery.Date.,
         Expense=Cost.Incurred...., Profit=Profit.Amount.....,Payment_details=Payment.details) 
```



### Categorize products based on Product type and quantity  


```{r}
Amoree2022_v2 <- add_column(Amoree2022_v2, Category=NA)


for (i in 1:nrow(Amoree2022_v2)) 
{
    if(str_detect(Amoree2022_v2[i,3], ("Birthday")) == TRUE)
    {
      Amoree2022_v2[i,"Category"]="Birthday"
      
    }
  else
    
    if(str_detect(Amoree2022_v2[i,3],("Monogram")) == TRUE)
    {
  
  Amoree2022_v2[i,"Category"]="Monogram"
}

else
  
  if(str_detect(Amoree2022_v2[i,3],("Frame")) == TRUE)
  {
    
    Amoree2022_v2[i,"Category"]="Frames"

  }
else
  
  Amoree2022_v2[i,"Category"]="Others"
  
}



for(i in 1:nrow(Amoree2022_v2))
  if((str_detect(Amoree2022_v2[i,3],("Liquor"))) | (str_detect(Amoree2022_v2[i,3],("Men"))) | (str_detect(Amoree2022_v2[i,3],("Alcohol")))  == TRUE)
  {
    
    Amoree2022_v2[i,"Category"]="Men's hampers"
    
  }

                        ### Categorize the Products based on Quantity

for(i in 1:nrow(Amoree2022_v2))
{
  if(Amoree2022_v2[i,6]>5)
    Amoree2022_v2[i,"Category"]="Return gifts/Bulk"
    
}
```


    
### Get a glimpse of the data frame structure  
  
  

```{r}
glimpse(Amoree2022_v2)
```




### Change the data types to appropriate data types according to the column names.  
### Also split Date into months and years for convenience.  


```{r}
Amoree2022_v2 <- transform(Amoree2022_v2,Price = as.numeric(Price),Expense = as.numeric(Expense),Date=dmy(Date))

Amoree2022_v2 <- Amoree2022_v2 %>% mutate(Month=month(Date)) %>%
                                        mutate(Year=year(Date)) %>% arrange(Month)
```


### Profit analysis calculations  

```{r}
Rev_2022 <- Amoree2022_v2 %>% group_by(Month) %>%  summarise(Revenue = sum(Price,na.rm=TRUE),
                                   Cost = sum(Expense,na.rm=TRUE),
                                   Profit = sum(Profit,na.rm=TRUE))

Profit_analysis_2022 <- Amoree2022_v2  %>% 
  select(Product,Qty,Profit,Price,Category) %>% mutate(Profit_per_item=Profit/Qty) %>% 
  group_by(Category) %>% 
  summarise(Avg_profit_peritem=mean(Profit_per_item, na.rm = TRUE))
View(Profit_analysis_2022)

Revenue_per_category_2022 <- Amoree2022_v2 %>% 
  select(Product,Qty,Profit,Price,Category) %>% group_by(Category) %>% 
summarise(Revenue_per_category=sum(Price,na.rm=TRUE),
          Total_qty_per_category=sum(Qty,na.rm=TRUE))

Profit_per_category_2022 <- Amoree2022_v2  %>% 
  select(Product,Qty,Profit,Price,Category) %>% group_by(Category) %>% 
  summarise(Profit_per_category=sum(Profit,na.rm=TRUE))


Profit_percentage_per_category_2022 <- Profit_per_category_2022 %>%
  mutate(Revenue_per_category=Revenue_per_category_2022$Revenue_per_category,
        Qty=Revenue_per_category_2022$Total_qty_per_category) %>% 
                mutate(Avg_price_per_category=Revenue_per_category/Qty,
                        Profit_percent_per_category=(Profit_per_category/Revenue_per_category)*100)

 View(Profit_percentage_per_category_2022)
```




## Plots



```{r , echo=FALSE}


ggplot(Amoree2022_v2)+ geom_bar(mapping=aes(x=Category))+ 
  labs(title="Number of orders by Category 2022")

ggplot(Rev_2022,aes(x=Month,y=Revenue)) + geom_point() + geom_line() +
  scale_x_continuous(breaks=seq(0,12, by=1)) +
     labs(title = "Revenue by months 2022",y="Revenue")

ggplot(Amoree2022_v2,aes(x=Month,y=Profit,fill=Category)) + 
  geom_col() + scale_x_continuous(breaks=seq(0,12.5,by=1)) +
  labs(title = "Profit by months 2022")

ggplot(Amoree2022_v2)+ geom_col(mapping=aes(x=Category,y=Price),fill="darkblue")+
  labs(title="Revenue per Category 2022",y="Revenue") 

ggplot(Amoree2022_v2)+ geom_col(mapping=aes(x=Category,y=Profit),fill="blue") +
  labs(title="Profit per Category 2022")

```

